---
title: "Midterm Project - Analysis of Cryptocurrency"
output:
  md_document:
    variant: markdown_github
---

```{r, echo=FALSE, warning = FALSE}
library(httr)
library(xml2)
library(stringr)
library(ggplot2)
library(dplyr)
```

# Introduction

Cryptocurrency is a digital or virtual currency that uses cryptography for security and operates independently of a central bank. Cryptocurrencies are decentralized, meaning that they are not controlled by a single authority or institution, but instead operate on a distributed ledger system called blockchain.

The first and most well-known cryptocurrency is Bitcoin, which was created in 2009 by an anonymous person or group of people using the pseudonym Satoshi Nakamoto. Since then, many other cryptocurrencies have been developed, including Ethereum, Litecoin, and Ripple.

Cryptocurrencies are created through a process called mining, which involves using computational power to solve complex mathematical problems and verify transactions on the blockchain. Transactions on the blockchain are recorded in blocks, which are then added to the blockchain in a chronological and immutable manner.

Now, cryptocurrency has become an increasingly important investment in recent years, with many investors and traders looking to capitalize on its price volatility and potential for high returns. However cryptocurrencies, like any other investment, carry a number of risks that investors should be aware of before investing. For example, cryptocurrencies are known for their high volatility, with prices often fluctuating significantly over short periods of time and the value of cryptocurrencies depends on their adoption and use by individuals and businesses. If adoption is low or slows down, the value of a cryptocurrency may decrease or become worthless. Thus while they can provide opportunities for high returns, investors should carefully consider timing of buying or selling cryptocurrencies.

In this project, we will explore the historical price data of some cryptocurrencies and try to gain insights into its price trends and factors that may affect its price. We will use various statistical and data visualization techniques to analyze the data and draw conclusions about the cryptocurrency's past and potential future performance.

To do this, we will collect general data and history price of cryptocurrencies by Coinranking API and use various statistical and data visualization techniques to analyze the data and draw conclusions about the cryptocurrency's past and potential future performance.

# Collect general data by API

At first, we need to extract general data including ids, names etc.
```{r, warning = FALSE}
url <- "https://coinranking1.p.rapidapi.com/coins"

coin_query_string <- list(
  referenceCurrencyUuid = "yhjMzLPhuIDl",
  timePeriod = "24h",
  orderBy = "marketCap",
  orderDirection = "desc",
  limit = "50",
  offset = "0"
)

query_coins <- VERB("GET", url, add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1', 'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'), query = coin_query_string, content_type("application/octet-stream"))

coins <- httr::content(query_coins)
coins <- as.character(coins)
```

## Extract features

The text data get from Coinranking contain lots of useless information in our task, for example icon url, btcPrice (Price of the coin expressed in Bitcoin), etc. Thus we just extract some variables which may affect the price of cryptocurrency.

```{r, warning = FALSE}
ids <- stringr::str_extract_all(coins, "uuid = \\\"[A-Za-z0-9_-]+\\\"")[[2]]
ids <- stringr::str_remove_all(ids, "uuid = \\\"|\\\"")

symbols <- stringr::str_extract_all(coins, "symbol = \\\"[A-Za-z0-9]+\\\"")[[2]]
symbols <- stringr::str_remove_all(symbols, "symbol = \\\"|\\\"")

names <- stringr::str_extract_all(coins, "name = \\\"[A-Za-z0-9 .()]+\\\"")[[2]]
names <- stringr::str_remove_all(names, "name = \\\"|\\\"")

colors <- stringr::str_extract_all(coins, "color = (\\\"[#A-Za-z0-9]+\\\")|NULL")[[2]]
colors <- stringr::str_remove_all(colors, "color = \\\"|\\\"")

tiers <- stringr::str_extract_all(coins, "tier = [0-9]")[[2]]
tiers <- stringr::str_remove_all(tiers, "tier = ")

prices <- stringr::str_extract_all(coins, "price = (\\\"[0-9.]+\\\")")[[2]]
prices <- as.numeric(stringr::str_remove_all(prices, "price = \\\"|\\\""))

volume24h <- stringr::str_extract_all(coins, "`24hVolume` = (\\\"[0-9.]+\\\")")[[2]]
volume24h <- as.numeric(stringr::str_remove_all(volume24h, "`24hVolume` = \\\"|\\\""))

changes <- stringr::str_extract_all(coins, "change = (\\\"-?[0-9.]+\\\")")[[2]]
changes <- as.numeric(stringr::str_remove_all(changes, "change = \\\"|\\\""))

marketCap <- stringr::str_extract_all(coins, "marketCap = (\\\"[0-9.]+\\\")")[[2]]
marketCap <- as.numeric(stringr::str_remove_all(marketCap, "marketCap = \\\"|\\\""))
```

## Form a database

```{r build-db, warning = FALSE}
database <- data.frame(
  ids, symbols, names, colors, tiers, prices, volume24h, changes, marketCap
)
```

Table 1, which is shown below contain basic information for first five cryptocurrencies by Coinranking API. Noticed that all price are calculated by US Dollar and rows in this table are sorted in descending order by marketCap (market capital).

```{r, warning = FALSE}
knitr::kable(head(database), caption = "Table 1: General Data of Some Cryptocurrency")
```


## Get other variables

In addition to the variables mentioned before, there are a number of other variables that may affect the price of cryptocurrency, like maximum amount the coin supply can ever be, current amount of coins that are in existence, etc. We need to use another API to get such data. We can also replace missing identifier with $NA$ in this step.

```{r, warning = FALSE}
max_amount <- c()
current_amount <- c()
circulatings <- c()
number_of_markets <- c()
number_of_exchanges <- c()

for (id in database$ids) {
  supply_url <- paste("https://coinranking1.p.rapidapi.com/coin/", id, sep = "")
  supply <- VERB("GET",
                 supply_url, 
                 add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                             'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                 content_type("application/octet-stream")
                 )
  supply <- httr::content(supply)
  supply <- as.character(supply)
  max_amount_id <- stringr::str_extract_all(supply, "max = \\\"[0-9.]+\\\"|max = NULL")[[2]]
  max_amount_id <- as.numeric(stringr::str_remove_all(max_amount_id, "max = |\\\"|\\\""))
  current_amount_id <- stringr::str_extract_all(supply, "total = \\\"[0-9.]+\\\"|total = NULL")[[2]]
  current_amount_id <- as.numeric(stringr::str_remove_all(current_amount_id, "total = |\\\"|\\\""))
  circulating <- stringr::str_extract_all(supply, "circulating = \\\"[0-9.]+\\\"|circulating = NULL")[[2]]
  circulating <- as.numeric(stringr::str_remove_all(circulating, "circulating = |\\\"|\\\""))
  market_number <- stringr::str_extract_all(supply, "numberOfMarkets = [0-9.]+|numberOfMarkets = NULL")[[2]]
  market_number <- as.numeric(stringr::str_remove_all(market_number, "numberOfMarkets = "))
  exchanges_number <- stringr::str_extract_all(supply, "numberOfExchanges = [0-9.]+|numberOfExchanges = NULL")[[2]]
  exchanges_number <- as.numeric(stringr::str_remove_all(exchanges_number, "numberOfExchanges = "))
  max_amount <- c(max_amount, max_amount_id)
  current_amount <- c(current_amount, current_amount_id)
  circulatings <- c(circulatings, circulating)
  number_of_markets <- c(number_of_markets, market_number)
  number_of_exchanges <- c(number_of_exchanges, exchanges_number)
}
```
## Advanced database with more variables

```{r, warning = FALSE}
database$max <- max_amount
database$current <- current_amount
database$circulatings <- circulatings
database$market_number <- number_of_markets
database$exchange_number <- number_of_exchanges
database$frequency_exchange <- ifelse(database$volume24h > median(database$volume24h), "high", "low")
database <- database %>% mutate(`V/MC` = volume24h/marketCap)
```

Table 2, which is shown below, contain more useful information for first five cryptocurrencies. In addition, we add another variables $frequency\_exchange$ and $V/MC$ based on volume in last 24 hour. Because a high volume and volume / market cap ratio means high liquidity. Simply put, that means a cryptocurrency can be easily bought or sold on an exchange, close to its value. This usually results in a more stable market with few fluctuations in price.

```{r, warning = FALSE}
knitr::kable(head(database), caption = "Table 2: More Data of Some Cryptocurrency")
```


# General Data Exploration

```{r, warning = FALSE}
table3 <- database %>%
  group_by(frequency_exchange) %>%
  summarise(median_price = median(prices), max_price = max(prices), min_price = min(prices), median_exchange_number = median(exchange_number),mean_exchange_number = mean(exchange_number), mean_ratio = mean(`V/MC`))
```

Table 3, which is shown below, show that generally speaking, cryptocurrency with high volume have higher expected number of exchange and significantly higher $V/MC$ ratio, which means a more stable market with few fluctuations in price. Thus these cryptocurrencies may be better for long-term holding than the rest.

```{r, warning = FALSE}
knitr::kable(table3, caption = "Table 3: Summary of Cryptocurrency group by volume")
```


From Graph 1 below, we can see that there is no correlation between last 24 hour changes and volume. In addition, most price of cryptocurrencies are falling recently. 

** But we I rerun code in March 10th, half of these cryptocurrencies are up, even one of them is up 6.55% in last 24 hour! I guess this was due to a speech about cryptocurrencies held by federal reserve system in yesterday or maybe this is just another proof of the huge price volatility of cryptocurrencies. I am not quite sure.

```{r, warning = FALSE}
ggplot(data = database, aes(reorder(changes, changes))) +
  theme_minimal() +
  geom_col(aes(y=changes, fill = frequency_exchange), position = "dodge") +
  labs(x=NULL, y=NULL, title = "Garph 1: 24 hour Changes of Cryptocurrency") +
  coord_flip()
```

# History price analysis

## Choose Some Cryptocurrencies

```{r, warning = FALSE}
max_change <- database[which.max(database$changes), ]$ids
max_volume <- database[which.max(database$volume24h), ]$ids
max_price <- database[which.max(database$prices), ]$ids
min_change <- database[which.min(database$changes), ]$ids
min_volume <- database[which.min(database$volume24h), ]$ids
min_price <- database[which.min(database$prices), ]$ids
```

In this case, we choose 6 cryptocurrencies by change, volume and price in last 24 hours.

## Extract history price of selected Cryptocurrency


```{r, warning = FALSE}
query_string <- list(timePeriod = "5y") # Extract the last five years data
```

In this case, we extract past 5 years price of these cryptocurrencies.

```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", max_change, "/history", sep="")
name <- database %>% filter(ids == max_change) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream")
                      )
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```
```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", max_volume, "/history", sep="")
name <- database %>% filter(ids == max_volume) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream"))
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```
```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", max_price, "/history", sep="")
name <- database %>% filter(ids == max_price) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream"))
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```
```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", min_change, "/history", sep="")
name <- database %>% filter(ids == min_change) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream"))
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```
```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", min_volume, "/history", sep="")
name <- database %>% filter(ids == min_volume) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream"))
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```
```{r, warning = FALSE}
price_url <- paste("https://coinranking1.p.rapidapi.com/coin/", min_price, "/history", sep="")
name <- database %>% filter(ids == min_price) %>% select(names) %>% pull()
history_price <- VERB("GET",
                      price_url, 
                      add_headers('X-RapidAPI-Key' = '9c9d129698mshb8275bce0446cbcp11a9c2jsn311a96af71b1',
                                  'X-RapidAPI-Host' = 'coinranking1.p.rapidapi.com'),
                      query = query_string,
                      content_type("application/octet-stream"))
history_price <- httr::content(history_price)
history_price <- as.character(history_price)
price <- stringr::str_extract_all(history_price, "price = (\\\"[0-9.]+\\\")|price = NULL")[[2]]
price <- stringr::str_remove_all(price, "price = |\\\"|\\\"")
timestamp <- stringr::str_extract_all(history_price, "timestamp = [0-9]+")[[2]]
timestamp <- stringr::str_remove_all(timestamp, "timestamp = ")
history_price <- data.frame(price, timestamp)
history_price$timestamp <- as.POSIXct(as.integer(history_price$timestamp), origin="1970-01-01", tz="UTC")
history_price$price <- as.numeric(history_price$price)
knitr::kable(head(history_price), caption = paste("history price of", name, sep = " "))
ggplot(history_price, aes(x=timestamp, y=price)) +
  geom_point() +
  geom_smooth() +
  labs(title = paste("history price of", name, sep = " "))
```


From all plot above, we can see that in the first six month of 2021, all cryptocurrency prices fall rapidly, after which most prices recover. But after 2022, cryptocurrency prices see an even steeper decline. 

Firstly, we may conclude that all cyptocurrencies are highly correlated. In the long term, they all have the same price trend. Thus different cyptocurrencies is not good investment assets, since it can amplify the risk of the portofolio. A decline in one asset's price can also lead to a decline in the price of the other asset. Additionally, based on Graph 1, I would say that this is not a good time to invest cryptocurrency, since price cryptocurrencies are highly correlated and most of them tend to decrease.

Secondly, through search engines, we can find that the price drop in 2021 could be due to the global economic downturn caused by the impact of the epidemic. Thus after that, the market of cryptocurrency saw a rebound. However, the second price drop was due to the collapse of FTX, a bankrupt company that formerly operated a cryptocurrency exchange and crypto hedge fund. I don't have much time to research more relevant information, but as far as I know, this is a black swan event. So it is almost unpredictable.

Noticed that, you may get different result when re-run the code, because the data always update.

# Conclusions

From the analysis above, we can get that cryptocurrencies are highly correlated, so invertors should not hold too many different cryptocurrencies, since it will arise risks of your assets. But maybe this is a good time to buy some cryptocurrencies, since price of most cryptocurrencies are slowly recovering. In addition, most of comment about bitcoin on social media, like twitter are positive. So maybe people's confidence in cryptocurrency is slowly rebounding as well and this may suggest that the future market for cryptocurrencies is optimistic. 

There are some limitations in this project. Firstly, this API doesn't provide historical supply, thus we can only try to observe the price trend by historical price. Secondly, there are lots of factors we did not consider in this project for now, such as, the price, quantity and computation resources of GPU (as I explain in introduction, one of approach to get cryptocurrencies is mining), policies of different countries on cryptocurrency and investor confidence in cryptocurrencies. In future, I may try to extract historical supply data or extract some comments about cryptocurrencies (for sentiment analysis, maybe generate a faith index on cryptocurrency by comment on social media) to do further reseacher and build model for predictions.